---
title: "Plots for catch by age from NetCDF output"
author: "Alberto Rovellini"
date: "6/16/2023"
output: html_document
---

```{r theme setup, include = F, echo=F}
knitr::opts_chunk$set(echo = FALSE,warnings=FALSE)
library(tidyverse)
library(sf)
library(rbgm)
library(tidync)
library(magrittr)
library(cowplot)
library(here)
library(viridis)
library(rnaturalearth)
library(reshape2)
library(RColorBrewer)
library(tictoc)
library(ggridges)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)

options(dplyr.summarise.inform = FALSE)
sf_use_s2(FALSE)
```

Output directory to check
```{r set paths, include=F, echo=F}
this_no <- '1275'
burn_in <- 8
#this_run <- '01'
this.dir <- paste0('C:/Users/Alberto Rovellini/Documents/GOA/Parametrization/output_files/data/out_', this_no)#, '_', this_run)
```

```{r spatial data,include=F,echo=F}
fl <- paste(this.dir,'GOA_WGS84_V4_final.bgm',sep='/')
# load the file
bgm <- read_bgm(fl)
# names(bgm)
goa_sf <- box_sf(bgm)
goa_sf <- goa_sf %>% mutate(box_id = box_id+1) #ALBI: adding this for correct matching of boxes
st_crs(goa_sf) <- st_crs(attr(goa_sf$geometry, "crs")$proj)

# load west cost land for mapping
coaststates <- ne_countries(continent="North America",returnclass = 'sf') %>% 
  filter(name %in% c('Canada','United States'))
```

```{r functional groups table,include=F,echo=F,message=F}
grps <- read_csv('../data/GOA_Groups.csv')

# set up a functional group types table
vertebrate_groups <- grps %>% filter(GroupType%in%c("FISH","SHARK","BIRD","MAMMAL")) %>% mutate(BiomassType="vertebrate")
plankton_groups <- grps %>% filter(GroupType %in% c("PWN",'CEP','LG_ZOO','MED_ZOO','SM_ZOO','LG_PHY','SM_PHY')) %>% 
  mutate(BiomassType="plankton")
bottom_groups <- grps %>% filter(GroupType %in% c("MOB_EP_OTHER",'SED_EP_FF','SED_EP_OTHER','PHYTOBEN')) %>% 
  mutate(BiomassType="2D")
other_groups <- grps %>% filter(GroupType %in% c("LG_INF","MICROPHTYBENTHOS","SED_BACT","PL_BACT","SM_INF","CARRION","LAB_DET","REF_DET"))%>% 
  mutate(BiomassType="other")
biomass_groups <- bind_rows(vertebrate_groups,plankton_groups,bottom_groups,other_groups)

# add to grps df
grps <- grps %>% left_join(biomass_groups)
```

```{r load atlantis output,include=F,echo=F}
tic("Loading files: ")
# output file
out_fl <- paste(this.dir, paste0('outputGOA', '0', this_no, '_test.nc'),sep='/')
out <- tidync(out_fl)
this_nc <- ncdf4::nc_open(out_fl)

# catch output file
out_fl_catch <- paste(this.dir, paste0('outputGOA', '0', this_no, '_testCATCH.nc'),sep='/')
out_catch <- tidync(out_fl_catch)
this_nc_catch <- ncdf4::nc_open(out_fl_catch)

# derived values for output
# depths <- out %>% hyper_filter(t=t==0) %>% hyper_tibble(select_var="dz") %>% dplyr::select(-t)
# glimpse(depths)

# volumes of each layer
volumes <- out %>% hyper_filter(t=t==0) %>% hyper_tibble(select_var="volume") %>% dplyr::select(-t)

# time dimension
ts <- ncdf4::ncvar_get(this_nc,varid = "t") %>% as.numeric
tyrs <- ts/(60*60*24*365)

# area of each box is the same as volume of the deepest depth layer, because the dz of that layer is 1
areas <- volumes %>% filter(z==max(z)) %>% dplyr::select(b,volume) %>% rename(area=volume)

# functional group dimensions
# fg_dimensions <- hyper_grids(out) %>% 
#   pluck("grid") %>% 
#   purrr::map_df(function(x){
#     out %>% activate(x) %>% hyper_vars() %>% 
#       mutate(grd=x)
#   })
```

```{r load harvest input,include=F,echo=F}
# read in harvest.prm parameter file
harvest_prm <- readLines(paste(this.dir, 'GOA_harvest_background.prm',sep='/'))

# get table with input CatchTS_agedistribXXX values
parnames <- harvest_prm[grep('CatchTS_agedistrib', harvest_prm)]
parvals <- harvest_prm[grep('CatchTS_agedistrib', harvest_prm)+1]
agedistrib <- strsplit(parvals, split = ' ') %>%
  lapply(.,function(x){x %>% as.numeric() %>% matrix(nrow=1) %>% data.frame()}) %>%
  bind_rows() %>%
  set_names(1:10) %>%
  mutate(Code = gsub(' ','',gsub(' .*','',gsub('CatchTS_agedistrib','',parnames)))) %>%
  pivot_longer(-Code, names_to = 'age', values_to = 'prop') %>%
  left_join(grps %>% dplyr::select(Code, Name)) 
```

Important function to remove the boundary box values from these calculations

```{r boundary,include=F,echo=F}
boundary_boxes <- goa_sf %>% filter(boundary == TRUE) %>% pull(box_id)

setNA <- function(mat) {
  mat2 <- mat
  if(length(dim(mat2))==3) mat2[,boundary_boxes,]<-NA
  if(length(dim(mat2))==2) mat2[boundary_boxes,] <- NA
  mat2
}
```

```{r catch fxn,include=F,echo=F}
# for catch
plot_catch <- function(fg,type){
  
  # if the species is TURNED OFF, return an empty plot
  if(grps$IsTurnedOn[grps$Name==fg]==0) return(tibble(t=0,catch=0) %>% ggplot(aes(t,catch))+geom_blank())
  # get the attributes associated with each functional group
  fg_atts <- grps %>% filter(Name==fg)
  #Extract from the output .nc file the appropriate time series variables
  catch_vars <- hyper_vars(out_catch %>% activate('D1,D0')) %>% # all variables in the .nc file active grid
    filter(grepl("_Catch",name)) %>% # filter for catch variables - these are numbers by age
    filter(grepl(fg,name)) # filter for specific functional group
  # Actually pull the data from the .nc
  catch1 <- purrr::map(catch_vars$name,ncdf4::ncvar_get,nc=this_nc_catch) %>% 
    lapply(setNA)
  
  catch2 <- list()
  
  for(i in 1:length(catch1)){
    
    this_age_group <- catch1[[i]]
    
    this_age_group <- this_age_group %>%
      data.frame() %>%
      set_names(0:40) %>%
      summarise(across(everything(), ~ sum(.x, na.rm = TRUE))) %>%
      #mutate(box_id = 0:(nrow(.)-1)) %>%
      pivot_longer(cols = everything(), names_to = 't', values_to = 'catch') %>%
      mutate(t = as.numeric(t),
             age_group = catch_vars$name[i],
             age = parse_number(age_group))
    
    catch2[[i]] <- this_age_group
  }
  
  catch3 <- catch2 %>% bind_rows()
  
  # filter time
  catch3 <- catch3 #%>%
  #filter(t < max(tyrs))#, t > burn_in) # drop last year as it often is not fully recorded?
  
  ###########################################################################################
  # now bring in weight at age, because catch3 contains numbers and we want catch in biomass
  #Extract from the output .nc file the appropriate reserve N time series variables
  resN_vars <- hyper_vars(out) %>% # all variables in the .nc file active grid
    filter(grepl("_ResN",name)) %>% # filter for abundance variables
    filter(grepl(fg,name)) # filter for specific functional group
  
  #Extract from the output .nc file the appropriate structural N time series variables
  strucN_vars <- hyper_vars(out) %>% # all variables in the .nc file active grid
    filter(grepl("_StructN",name)) %>% # filter for abundance variables
    filter(grepl(fg,name)) # filter for specific functional group
  
  # Get numbers by box
  abun_vars <- hyper_vars(out) %>% # all variables in the .nc file active grid
    filter(grepl("_Nums",name)) %>% # filter for abundance variables
    filter(grepl(fg,name)) # filter for specific functional group
  
  if(nrow(resN_vars)==0) {return("no data.")}
  else {
    # # Actually pull the data from the .nc
    resN <- purrr::map(resN_vars$name,ncdf4::ncvar_get,nc=this_nc) 
    strucN <- purrr::map(strucN_vars$name,ncdf4::ncvar_get,nc=this_nc)
    nums <-purrr::map(abun_vars$name,ncdf4::ncvar_get,nc=this_nc) #numbers by age group,box,layer,time
    totnums <-nums %>% purrr::map(apply,MARGIN=3,FUN=sum) # total numbers by age group, time
    relnums <- purrr::map2(nums,totnums,sweep,MARGIN=3,FUN=`/`) # divide nums by totnums along the time axis to get relative annual nums per age group/box/layer
    
    # add the two matrices to get total nitrogen weight
    rnsn <- purrr::map2(resN,strucN,`+`)
    
    # multiply and sum to get abundance-weighted mean weight at age
    rnsn_summ <- purrr::map2(rnsn,relnums,`*`) %>% 
      purrr::map(apply,MARGIN=3,FUN=sum) %>% # mean total N by time
      bind_cols() %>% # bind age groups elements together
      suppressMessages() %>% 
      set_names(resN_vars$name) %>% 
      mutate(t=tyrs) %>%
      # pivot to long form
      pivot_longer(cols = contains(fg_atts$Name),names_to = 'age_group',values_to = 'totN') %>%
      mutate(age=parse_number(age_group)) %>% 
      mutate(weight=totN*20*5.7/1000000) %>%   # convert totN to weight/individual in kg
      dplyr::filter(t %in% 0:max(tyrs)) # this is because the catch netcdf is not written out as frequently
    
  }
  
  ###########################################################################################
  # add in numbers per time step
  abun1 <- purrr::map(abun_vars$name,ncdf4::ncvar_get,nc=this_nc) %>% 
    lapply(setNA) %>%
    purrr::map(apply,MARGIN=3,FUN=sum,na.rm=T) %>% 
    bind_cols() %>% 
    suppressMessages() %>% 
    set_names(abun_vars$name) %>% 
    mutate(t=tyrs) %>%
    dplyr::filter(t %in% 0:max(tyrs)) # this is because the catch netcdf is not written out as frequently
  
  abun2 <- abun1 %>%
    pivot_longer(cols = contains(fg_atts$Name),names_to = 'age_group',values_to = 'abun') %>%
    mutate(age=parse_number(age_group))
  
  ###########################################################################################
  # now join catch in numbers with the weight at age over time and by age class and total numbers
  catch4 <- catch3 %>%
    left_join(rnsn_summ %>% dplyr::select(t, age, weight)) %>%
    left_join(abun2 %>% dplyr::select(t, age, abun)) %>%
    mutate(catch_mt = catch * weight / 1000, # go from kg to t
           selex = catch / abun) %>%# get selectivity in numbers
    rowwise() %>%
    mutate(selex = ifelse(selex>1,NA,selex)) %>% # if more individuals are caught than there are, it means they have gone / are going exctinct
    ungroup()
  
  # make a different plot depending on the type argument
  
  if(type == 'ts'){
    
    # plot
    plot_out <- catch4 %>%
      filter(t>0) %>% # remove plotting artifact
      ggplot(aes(t,catch_mt,col=factor(age)))+
      geom_line()+
      geom_vline(xintercept = burn_in, color = 'red', linetype = 'dashed')+
      labs(col="Age Group",y="Catch (mt)",x="Year",title=paste0(fg_atts$LongName,"\nCatch"))
    
  } else if (type == 'prop') {
    
    this_agedistrib <- agedistrib %>%
      filter(Name == fg) %>%
      mutate(t = burn_in, type = 'input', age = as.numeric(age)) %>%
      dplyr::select(type, t, age, prop)
    
    plot_out <- catch4 %>%
      filter(t>0) %>% # remove plotting artifact
      group_by(t) %>%
      mutate(tot_catch = sum(catch_mt)) %>%
      ungroup() %>%
      mutate(type = 'output', prop = catch_mt / tot_catch) %>%
      dplyr::select(type,t,age,prop) %>%
      rbind(this_agedistrib) %>%
      ggplot(aes(t,prop,fill=factor(age)))+
      geom_bar(stat = 'identity', position = 'stack')+
      geom_vline(xintercept = burn_in, color = 'red', linetype = 'dashed')+
      labs(fill="Age Group",y="Proportion of total catch",x="Year",title=paste0(fg_atts$LongName,"\nCatchTS_agedistribXXX"))+
      facet_grid(~type, scales = 'free_x', space = 'free_x')
    
  } else if (type == 'selex') {
    
    plot_out <- catch4 %>%
      filter(t %in% seq(min(t), max(t), 8)) %>% # don't need one per year, thin them out
      filter(t < max(tyrs)) %>%
      ggplot(aes(x = age, y = selex))+
      geom_line(linewidth = 1)+
      theme_bw()+
      scale_x_continuous(breaks = 1:10)+
      #scale_y_continuous(breaks = seq(0,1,0.2), limits = c(0,1))+
      labs(y="Proportion caught",x="Age class",title=paste0(fg_atts$LongName," selectivity"))+
      facet_grid(~t)
    
  } else {
    
    stop('This type does not exist')
    
  }
  
  return(plot_out)
}
```

```{r calc plots,warning=FALSE,message=F}
# Here we calculate and store the actual plots
# catch ts
tic("Catch time series plots: ")
verts <- grps %>% filter(GroupType%in%c("FISH","SHARK","BIRD","MAMMAL")) %>% pluck('Name')
verts_catch_at_a <- purrr::map(verts,~plot_catch(., type = 'ts'))
verts_names <- grps %>% filter(GroupType%in%c("FISH","SHARK","BIRD","MAMMAL")) %>% pluck('LongName')
toc()

# catch prop
tic("Proportion of caught biomass per age class plots: ")
verts_fraction_catch_by_a <- purrr::map(verts,~plot_catch(., type = 'prop'))
toc()

# selex
tic("Proportion of caught individuals plots: ")
verts_prop_caught_ind_at_a <- purrr::map(verts,~plot_catch(., type = 'selex'))
toc()
```

# Catch timeseries {.tabset .tabset-pills}

```{r ts,fig.height=6,fig.width=8,results='asis'}
if(length(verts_catch_at_a)>0){
  for(i in 1:length(verts_catch_at_a)){
    cat("  \n##",  verts_names[i],"  \n")
    print(verts_catch_at_a[[i]]) 
    cat("  \n")
  }
}
```

# Fraction of catch per age class - biomass {.tabset .tabset-pills}

Vertical dashed line indicates when catch.ts kicks in. Panel labeled 'input' with one bar shows the proportion of catch as specified in the parameter `CatchTS_agedistribXXX1` in harvest.prm. 

```{r agedistrib,fig.height=6,fig.width=8,results='asis'}
if(length(verts_fraction_catch_by_a)>0){
  for(i in 1:length(verts_fraction_catch_by_a)){
    cat("  \n##",  verts_names[i],"  \n")
    print(verts_fraction_catch_by_a[[i]]) 
    cat("  \n")
  }
}
```

# Proportion of caught individuals by age class - numbers {.tabset .tabset-pills}

Panels represent simulation years. Last year not shown.  

```{r selex,fig.height=6,fig.width=8,results='asis'}
if(length(verts_prop_caught_ind_at_a)>0){
  for(i in 1:length(verts_prop_caught_ind_at_a)){
    cat("  \n##",  verts_names[i],"  \n")
    print(verts_prop_caught_ind_at_a[[i]]) 
    cat("  \n")
  }
}
```